---
title: "Lab 1"
author: "Robert Cervantes"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

These are the tasks for lab 1


## Task 1
Obtain the working directory

```{r}
getwd()
```


## Task 2 
First we read in the data given and examine the first 6 lines of the data in DDT.

```{r}
ddt <- read.csv("DDT.csv")

# First six lines
head(ddt)
```

## Task 3 
- The qualitative variables in DDT are SPECIES and RIVER.

- The quantitative variables in DDT are MILE, LENGTH, WEIGHT, and DDT.

```{r}
unique_species <- unique(ddt$SPECIES)
unique_species
```

- There are 3 species in DDT as shown above with the provided code^

```{r}
with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",])
```
- With this code we can subset the data to see instances where SPECIES = LMBASS and WEIGHT > 800gms.

```{r}
with(ddt, ddt[RIVER == "SCM" & DDT > 4.0,])
```
- With this code we can subset the data to see instances where RIVER = SCM and DDt > 4.0

## Task 4 
Q1. Using 
```{r}
summary(ddt)
```
we can see the mean length of fish is 42.81

Q2. Using 
```{r}
sd(ddt$WEIGHT)
```
we can see the standard deviation of weight is 376.5461

Q3. 0 

Q4.Using 
```{r}
v=1:20
v/20
```
we see that the last value is 1.00

## Task 5 
- Table of rivers 
```{r}
river_table <- table(ddt$RIVER)
river_table
```

- Color Coded barplot of rivers
```{r}
barplot(river_table, 
        col = rainbow(length(river_table)),
        xlab = "Rivers",
        ylab = "Number of Fish")
```

- Table of rivers crossed with species of fish
```{r}
crossed_table <- table(ddt$RIVER, ddt$SPECIES)
crossed_table
```

- Barlplot of rivers crossed with species of fish
```{r}
barplot(crossed_table, 
        legend.text = TRUE,
        xlab = "Rivers",
        ylab = "Number of fish")
```

## Task 6

- Pie chart of fish species
```{r}
species_table <- table(ddt$SPECIES)
pie(species_table)
```
Each slice represents how many fish are of that specific species


- Pie chart of rivers
```{r}
pie(river_table) ##we had a river table already!
```
Each slice shows a river and the size of slice corresponds to the number of fish from that river


## Task 7

- BoxPlot of the variable DDT 
```{r}
boxplot(ddt$DDT)
```

- Boxplot of the variable weight 
```{r}
boxplot(ddt$WEIGHT)
```

- Boxplot of the variable length
```{r}
boxplot(ddt$LENGTH)
```

## Task 8

coplot of LENGTH Vs WEIGHT given RIVER 
```{r}
coplot(LENGTH ~ WEIGHT | RIVER, data = ddt)
```

coplot of DDT Vs WEIGHT given SPECIES 
```{r}
coplot(DDT ~ WEIGHT | SPECIES, data = ddt)
```

## Task 9
First load package
```{r}
library(ggplot2)
```


- ggplot #1
```{r}
ggplot(ddt, aes(x = SPECIES, y = WEIGHT, fill = RIVER)) +
  geom_boxplot() +
  ggtitle("Robert Cervantes") +
  xlab("SPECIES") +
  ylab("WEIGHT") 
```

- ggplot #2
```{r}
ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES)) +
  geom_violin() +
  ggtitle("Robert Cervantes") +
  xlab("RIVER") +
  ylab("LENGTH")
```

- ggplot #3
```{r}
ggplot(ddt, aes(x = WEIGHT, y = LENGTH, color = SPECIES)) +
  geom_point() +
  ggtitle("Robert Cervantes") +
  xlab("RIVER") +
  ylab("LENGTH") 

```

















